Defining properties regulating pioneer-factor binding and function

### 2022-12-07 manuscript/manuscript.qmd
update MS results

-   Define binding classes I-III and describe characteristics (distribution, promoters/enhancers, etc.)

@@

To characterize the extent of chromatin opening by Zld and Grh, we integrated our ChIP-seq and ATAC-seq data and identified 3 classes of binding sites (Fig. 1a,b). Class I sites are accessible in WT cells and remain accessible when bound by ectopically-expressed Zld or Grh. Class II sites are inaccessible in WT cells, however binding of Zld or Grh does not lead to chromatin opening. Class III sites are inaccessible in WT cells and Zld or Grh binding results in chromatin opening. The majority of ChIP-seq peaks for Zld and Grh were class I sites, with a subset of binding sites in closed chromatin, and a small subset undergoing chromatin opening (Fig. 1c,f). While class I binding sites occured mostly at promoters, the majority of class II and III sites were at promoter-distal sites (Fig. 1d,g). Genome-wide analysis of signal intensities at all class II and III sites confirmed that these sites lack ATAC signal in WT cells and that class III sites undergo robust chromatin opening upon Zld or Grh induction (Fig. 1e,h). Together, these data show that when expressed exogenously at physiolgical levels, Zld and Grh can bind and open closed chromatin at a subset of binding sites.

@@

>>>>

-   ~~Define binding classes I-III and describe characteristics (distribution, promoters/enhancers, etc.)~~

@@

To characterize the extent of chromatin opening by Zld and Grh, we integrated our ChIP-seq and ATAC-seq data and identified 3 classes of binding sites (Fig. 1a,b). Class I sites are accessible in WT cells and remain accessible when bound by ectopically-expressed Zld or Grh. Class II sites are inaccessible in WT cells, however binding of Zld or Grh does not lead to chromatin opening. Class III sites are inaccessible in WT cells and Zld or Grh binding results in chromatin opening. The majority of ChIP-seq peaks for Zld and Grh were class I sites, with a subset of binding sites in closed chromatin, and a small subset undergoing chromatin opening (Fig. 1c,f). While class I binding sites occured mostly at promoters, the majority of class II and III sites were at promoter-distal sites (Fig. 1d,g). Genome-wide analysis of signa <!--#  -->l intensities at all class II and III sites confirmed that these sites lack ATAC signal in WT cells and that class III sites undergo robust chromatin opening upon Zld or Grh induction (Fig. 1e,h). Together, these data show that when expressed exogenously at physiolgical levels, Zld and Grh can bind and open closed chromatin at a subset of binding sites.

<!--# Could potentially add the cut&run time course here? Think about where to put this, or if it is worth including at all. -->

@@

Having established that ectopically expressed Zld and Grh bind and open closed chromatin, we sought to determine if binding of these factors also led to transcriptional activation. To identify changes in gene expression upon Zld or Grh induction, we performed RNA-seq and compared induced cells to WT cells lacking Zld or Grh expression. Differential expression analysis using DESeq2 [@love2014] identified XXX differentially expressed genes for Zld and XXX for Grh (Extended Data Fig. XX).
